---
title: "datelife R Package Reports"
date: "`r format(Sys.time(), '%d %B %Y')`"
author: Luna L. Sanchez Reyes
output: pdf_document
fontsize: 12pts
textalign: left
geometry: margin = 1in
---
```{r setup, include = FALSE, warnings = FALSE, message = FALSE}
getwd()
loadd(aves_spp)
loadd(aves_dr)
loadd(aves_mat)
loadd(aves_phylo)

loadd(aves400.1.gfr.runtime_2017.12.28)
loadd(aves.all.gfr.runtime_2017.12.29)
loadd(aves.all.gfr.runtime_2018.01.12)
loadd(res28)
loadd(res29)
loadd(res01)
loadd(res04)
```
# Benchmarking Functions to Generate New Data
Again, you'll need to install and load `datelife` and `microbenchmark` packages:
```{r eval = FALSE}
install.packages("microbenchmark")
install.packages("datelife")
library(microbenchmark)
library(datelife)
```

To benchmark `datelife`s functions for generation of new data on time of lineage divergence we chose birds as model.
First, we need to get all chronograms available. Using the following functions from `datelife` package:

```{r}
install.packages("datelife")
library(datelife)
aves_spp <- make_datelife_query(input="Aves", get_spp_from_taxon=TRUE)
# there are 12750 bird spp names
save(aves_spp, file="data/0_global/aves_spp.rda")
names(aves_spp)
length(aves_spp$cleaned_names)
aves_dr <- get_datelife_result(input = aves_spp$cleaned.names)
```
Check the structure of the object:
```{r}
length(aves_dr)
class(aves_dr)
unname(sapply(aves_dr, nrow))
# the largest tree comes from study:
names(aves_dr)[which.max(sapply(aves_dr, nrow))]
# and has more than 9k tips:
nrow(aves_dr[[which.max(sapply(aves_dr, nrow))]])
# save it
save(aves_dr, file="data/0_global/aves_dr.rda")
```
<!-- Now summarize results of the `datelife` search as a `phylo` object:
```{r}
start_time <- Sys.time() # to register run time
aves_phylo <- summarize_datelife_result(datelife_query = aves_spp$cleaned.names, datelife_result = aves_dr, summary_format = "phylo_all")
runtime <- Sys.time() - start_time # end of registering function running time
attr(aves_phylo, "runtime") <- runtime
save(aves_phylo, file = "data/0_global/aves_phylo.rda")
``` -->
<!-- Summarizing as phylo all bird chronograms took me `r paste(as.character(round(attr(aves_phylo, "runtime"), digits = 2)), attr(attr(aves_phylo, "runtime"), "units"))`
if the previous does not work, assign aves_phylo runtime to an object inside the plan. -->

Next step is to generate a sample of chronograms of different size that can be congruified and dated after.
These are going to be your target chronograms
```{r}
# it is much faster to drop tips from matrix:
tip_number <- c(10, 100, 200, 300, 400, 500, 600, 700, 800, 900, 1000, 2000, 3000, 4000, 5000, 6000, 7000, 8000, 9000)
aves_mat <- vector(mode = "list", length = length(tip_number))
length(aves_mat)
for(i in seq(tip_number)){
    print(i)
    taxa <- sample(seq(nrow(aves_dr[[which.max(sapply(aves_dr, nrow))]])), tip_number[i])
    aves_mat[[i]] <- aves_dr[[which.max(sapply(aves_dr, nrow))]][taxa, taxa]
}
sapply(aves_mat, nrow)
save(aves_mat, file="data/0_global/aves_mat.rda")
# remove large objects so your R session does not collapse
rm(aves_dr)
aves_trees <- summarize_datelife_result(datelife_result = aves_mat, summary_format = "phylo_all")
# this takes so much time, so its better to do it one by one, something like this:
# use datelife:::
plot(aves_trees)
class(aves_trees)
sapply(aves_trees, ape::is.ultrametric,  option = 2)
# they are all ultrametric
sapply(aves_trees, ape::is.binary.tree)
# and also fully dichotomous
for (i in seq(aves_trees)){
    xx <- aves_trees[[i]]
    assign(x = paste0("aves_tree_", tip_number[i]), value = xx)
    save(list = paste0("aves_tree_", tip_number[i]),
        file = paste0("data/0_global/aves_targets/aves_tree_", tip_number[i], ".rda"))
    rm(list = c(paste0("aves_tree_", tip_number[i]), "xx")) # to save space remove everything on your way out
}
# save(aves_trees, file="data/0_global/aves_trees.rda")
rm(aves_trees)
```
Now generate a 25%, 50% and 75% sample of subtrees for congruification

```{r}
# we will generate a lot of big matrices, so the best approach is to generate and
# save them each separately with a loop
# Actally, the function will have the same problem of saving all in cache
# so I finally went for a loop:
for(samp in c(0.25, 0.5, 0.75)){
    for (i in seq(length(aves_mat))){
        print(paste(samp, i))
        for (j in seq(100)){
            taxa <- sample(nrow(aves_mat[[i]]), ceiling(nrow(aves_mat[[i]])*samp))
            matname <- paste0("samp", samp*100, "_mat", i, "_", j)
            mm <- aves_mat[[i]][taxa,taxa]
            assign(x = matname, value = mm)
            save(list = matname, file = paste0("data/0_global/aves_mat_sample/", matname, ".rda"))
            rm(list = c("mm", matname))
        }
    }
}
```

Now congruify and date:

```{r}
for(samp in c(0.25, 0.5, 0.75)){
    gac_mbk <- uac_mbk <- data.frame()
    # trees <- rep(list(vector(mode = "list", length = 100)), length(tip_number))
    for (i in seq(length(tip_number))){
        phyname <- paste0("aves_tree_", tip_number[i])
        load(paste0("data/0_global/aves_targets/", phyname, ".rda"))
        phy <- get(phyname)
        for (j in 1:100){
            print(paste("samp", samp*100, "matrix", i, "rep", j))
            matname <- paste0("samp", samp*100, "_mat", i, "_", j)
            load(file = paste0("data/0_global/aves_mat/", matname, ".rda"))
            phylo <- patristic_matrix_to_phylo(patristic_matrix = get(matname))
            y <- microbenchmark(df <- get_all_calibrations(input= phylo), times=1L)
            levels(y$expr)[1] <- paste0("samp", samp*100, "_mat", i)
            gac_mbk <- rbind(gac_mbk, y)
            obj_name <- paste0(matname, "_gac")
            assign(value = df, x = obj_name)
            save(list = obj_name, file = paste0("data/2_datagen/results/", obj_name, ".rda"))
            rm(list = obj_name)
            y <- microbenchmark(tree <- use_all_calibrations(phy = phy, all_calibrations = df), times=1L)
            levels(y$expr)[1] <- paste0("samp", samp*100, "_mat", i)
            uac_mbk <- rbind(uac_mbk, y)
            obj_name <- paste0(matname, "_uac")
            assign(value = tree, x = obj_name)
            save(list = obj_name, file = paste0("data/2_datagen/results/", obj_name, ".rda"))
            rm(list = c(obj_name, matname, phyname))
            rm(df, tree)

            # save microbemchmark data in case the session crashes:
            obj_name <- paste0("samp", samp*100, "_gac_mbk")
            assign(value = gac_mbk, x = obj_name)
            save(list = obj_name, file = paste0("data/2_datagen/", obj_name, ".rda"))
            rm(list = obj_name)
            obj_name <- paste0("samp", samp*100, "_uac_mbk")
            assign(value = uac_mbk, x = obj_name)
            save(list = obj_name, file = paste0("data/2_datagen/", obj_name, ".rda"))
            rm(list = obj_name)
        }
        rm(list = phyname)
    }
}
source("data-raw/runtemp.R")
xx <- get_all_calibrations(input)
```
